---
title: "Earthquakes data exploration"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggmap)
library(ggplot2)
library(Hmisc)
URL <- "http://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv"
Earthquake_30Days <- read.table(URL, sep = ",", header = T)
```
## Data Exploration:
### lets check the head of the dataset
```{r echo=FALSE}
head(Earthquake_30Days)
```
as we can see we have 2 datetime columns 12 numeric columns and 8 textual columns  
one of the important things we can see is that we have missing values  

### lets explore some data statistics
```{r echo=FALSE}
describe(Earthquake_30Days)
```


## Data Description:

## Data Analysis:

## Map/s from the data:
```{r echo=FALSE,message=FALSE,warning=FALSE}
map <- get_map(location = c(lon = mean(Earthquake_30Days$longitude), lat = mean(Earthquake_30Days$latitude)), zoom = 3,maptype = "satellite", scale = 2)
ggmap(map) + geom_point(data = Earthquake_30Days, aes(x = longitude, color = mag,y = latitude, alpha = 0.1), size = 2, shape = 21) +  guides(fill=FALSE, alpha=FALSE, size=FALSE) + scale_color_gradient(low="yellow", high="red")
```

## Summary and Recommendations:


## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
