---
title: "Earthquakes data exploration"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggmap)
library(ggplot2)
library(Hmisc)
URL <- "http://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv"
Earthquake_30Days <- read.table(URL, sep = ",", header = T)
```
## Data Exploration:
### lets check the head of the dataset
```{r echo=FALSE}
head(Earthquake_30Days)
```
as we can see we have 2 datetime columns 12 numeric columns and 8 textual columns  
one of the important things we can see is that we have missing values  

### lets explore some data statistics
```{r echo=FALSE}
describe(Earthquake_30Days)
```
we can see there are some columns that contains many missing values such as 'nst', 'gap', 'dmin, 'magError' and 'magNst' 
i will omit those columns and remove rows with missing data on the rest of the data so i can have clean data to work with and also filter the non earthquake records
```{r echo=FALSE}
drops <- c("nst","gap","dmin","magError","magNst")
dataClean <- Earthquake_30Days[, !(names(Earthquake_30Days) %in% drops)]
dataClean <- dataClean[complete.cases(dataClean),]
dataClean <- dataClean[dataClean$type=="earthquake",]
```


## Data Description:
the data is about earthquakes telemetry which includes the location of the activity, the manitude, depth etc.  
as i see it the most relevant data attributes are the location represented by latitude and longtitude, the manitude, depth and rms(the root mean square of the travel time residual) and the location source.  
i will explore the relation of those attributes  

## Data Analysis:

### check correlation in the data columns

### graphs of the more "relevant" columns and compared to each other

### trend i see in the graphs

### try to do regression on some of the columns

## Map/s from the data:
```{r echo=FALSE,message=FALSE,warning=FALSE}
map <- get_map(location = c(lon = mean(Earthquake_30Days$longitude), lat = mean(Earthquake_30Days$latitude)), zoom = 3,maptype = "satellite", scale = 2)
ggmap(map) + geom_point(data = Earthquake_30Days, aes(x = longitude, color = mag,y = latitude, alpha = 0.1), size = 2, shape = 21) +  guides(fill=FALSE, alpha=FALSE, size=FALSE) + scale_color_gradient(low="yellow", high="red")
```
in the map we can see that most of the earthquakes are in the bounderies between continents and espacially in the "ring of fire" which is around the pacific ocean  

## Summary and Recommendations:
